---
title: "Script to plot individual figures"
author: "Christian Holland"
date: "6/23/2019"
output: html_document
---

```{r "knitr config", cache=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::knit(..., quiet = TRUE)
```
### Libraries and sources
These libraries and sources are used in this analysis 
```{r "setup", message=F}
library(tidyverse)
library(cowplot)
library(tidylog)
library(AachenColorPalette) # devtools::install_github("christianholland/AachenColorPalette")
library(ggpubr)
library(UpSetR)
library(VennDiagram)
library(lemon)
library(gridExtra)
library(lemon)
library(scales)
library(ggplotify)
library(pheatmap)
library(fgsea)


options("tidylog.display" = list(print))
theme_set(theme_cowplot())
source("src/my_ggplot_themes.R")

```
### Utility functions
```{r "utility functions"}
make_gsea_plot = function(data, geneset, time, zonation, ...) {
    stats_df = data %>%
      rename(stat = statistic) %>%
      arrange(-stat) %>% 
      mutate(stat = stat/max(abs(stat)),
             rank = row_number()) 
    
    rnk = stats_df$rank
    
    adj_stats = stats_df %>%
      select(gene, stat) %>%
      deframe()
  
    pathway = match(geneset, names(adj_stats)) %>%
      na.omit() %>%
      as.vector() %>%
      unname() %>%
      sort()
    
    
    gseaRes = calcGseaStat(adj_stats, selectedStats = pathway, returnAllExtremes = TRUE)
    bottoms = gseaRes$bottoms
    tops = gseaRes$tops
    n = length(adj_stats)
    xs = as.vector(rbind(pathway - 1, pathway))
    ys = as.vector(rbind(bottoms, tops))
    
    enrichment = tibble(rank = c(0, xs, n + 1), 
                        running_sum = c(0, ys, 0))
    ticks = tibble(rank = pathway, 
                   max_rank = max(rnk))
    color_df = enframe(rnk) %>%
      mutate(bin = factor(ntile(value, 10)))
    
    tmp = enrichment %>% 
      filter(abs(running_sum) == max(abs(running_sum)))
    
    g1 = ggplot(data = enrichment, aes(x = rank, y = running_sum)) + 
      geom_hline(yintercept = max(tops), colour = aachen_color("bordeaux"), linetype = "dashed") +
      geom_hline(yintercept = min(bottoms), colour =aachen_color("bordeaux"), linetype = "dashed") +
      geom_hline(yintercept = 0, colour = "black") +
      geom_segment(aes(x = tmp$rank, y = 0, xend = tmp$rank, yend = tmp$running_sum), 
                   linetype = "dashed") +
      geom_line(size=1, color = if_else(zonation == "periportal", aachen_color("turquoise"), aachen_color("violet"))) +
      lims(x = c(0, max(rnk) + 1)) +
      labs(y = "Enrichment score (ES)") +
      background_grid(major = "y", minor = "none", size.major = 0.4) +
      theme(axis.line.x = element_blank(),
            axis.title.x = element_blank(),
            axis.ticks.x = element_blank(),
            axis.text.x = element_blank(),
            plot.margin=unit(c(1,1,-0.25,1), "cm"),
            legend.position = "none",
            title = element_text(size=16),
            axis.text = element_text(size=14),
            plot.title = element_text(hjust = 0.5)
            #axis.line.y = element_blank()
            ) +
      annotate("text", x = c(1) , y = c(max(tops)+0.05), 
               label = c(str_c("ES =", round(gseaRes$res,3), sep = " ")),
               size=5, hjust = 0) +
      ggtitle(str_c("Month",time, "-", str_to_title(zonation), sep = " "))
    
    
    g2 = ggplot(color_df, aes(x=value, y=1)) +
      geom_tile(aes(color=value)) +
      geom_segment(ticks, mapping = aes(x = rank, y = 1.51, xend = rank, yend = 3.51), size = 0.5, color=aachen_color("black"), alpha=0.5) +
      scale_color_gradientn(colours = aachen_color(c("bordeaux", "orange", "yellow", "green", "blue", "violet"))) +
      scale_x_continuous(breaks = c(1,5000,10000,15000)) +
      theme(
          # axis.line.y = element_blank(),
          axis.ticks.y = element_blank(),
          axis.text.y = element_text(color="white"),
          axis.text.x = element_text(),
          plot.margin=unit(c(-0.25,1,1,1), "cm"),
          legend.position = "none",
          # axis.ticks.x = element_blank(),
          axis.line.x = element_blank(),
          title = element_text(size=16),
          axis.text = element_text(size=14)
            ) +
      labs(x = "Rank", y=NULL) 
    
    plot_grid(g1, g2, ncol = 1, align = "v", axis="l", rel_heights = c(2,1))
}


make_div_color_breaks = function(mat, low, mid, high) {
  paletteLength = 100
  colors = colorRampPalette(c(low, mid, high))(paletteLength)
  breaks = c(seq(min(mat), 0, length.out=ceiling(paletteLength/2) + 1),
            seq(max(mat)/paletteLength, max(mat), length.out=floor(paletteLength/2)))
  
  return(list(colors = colors, breaks = breaks))
}

```


### CCL4 experiment
#### Exploratory analysis
##### PCA
```{r "pca"}
pca = readRDS("output/ccl4_experiment/pca.rds")
explained = pca$explained
coords = pca$coords %>%
  mutate(dose = case_when(dose == "D0.6" ~ "0.6 ml/kg CCL4",
                          dose == "oil" ~ "Oil",
                          dose == "D0" ~ "Untreated")) %>%
  mutate(dose = factor(dose, levels = c("Untreated", "Oil", "0.6 ml/kg CCL4")),
         time = as.numeric(as.character(time))/4)

pca_all_1_2_text = ggplot(coords, aes(x=PC1, y=PC2, label=time, color=dose)) +
  geom_text(size=5, alpha = 0.8) +
  labs(x = str_c("PC1 (", explained[1], "%)"),
       y = str_c("PC2 (", explained[2], "%)"),
       color = "Treatment") +
  my_theme(grid = "no") +
  scale_color_manual(values = aachen_color(c("orange", "petrol", "red")), 
                     labels = c(expression("Untreated"), expression("Olive oil"), expression(CCl['4']*" + Olive oil"))) +
  theme(
    legend.position = "top",
    legend.text.align = 0,
    legend.box = "vertial",
    legend.title.align = 0.5,
    legend.box.just = "left")

saveRDS(pca_all_1_2_text, "fig/ccl4_experiment/pca_all_text.rds")
```

#### Differential gene expression analysis
##### Top 40 CCl4 genes
```{r "top 40 ccl4 genes"}
limma_result = readRDS("output/ccl4_experiment/limma_result.rds") %>%
  filter(contrast_reference == "ccl4") %>%
  mutate(time = str_extract(contrast, "\\d*$"),
         time = ordered(as.integer(time))) %>%
  mutate(label = fct_inorder(str_c("Week", time, sep=" "))) %>%
  mutate(regulation = str_to_title(regulation)) %>%
  mutate(regulation = factor(regulation, levels = c("Up", "Down", "No"))) %>%
  filter(regulation != "No") %>%
  group_by(time, regulation) %>%
  top_n(20, abs(logFC)) %>%
  ungroup()

up_8 = limma_result %>%
  filter(time == 8 & regulation == "Up") %>%
  ggplot(aes(x=fct_reorder(gene, logFC), y=logFC, fill=as_factor(sign(logFC)))) +
  geom_col() +
  labs(x=NULL, y="Up-regulated logFC") +
  scale_fill_manual(values = aachen_color("green")) +
  scale_y_continuous(position = "right", 
                     breaks = function(x) unique(floor(pretty(seq(0, (max(x) + 1) * 1.1))))) +
  theme(legend.position = "none",
        axis.text = element_text(size=10),
        axis.title = element_text(size=12)) +
  coord_flip()


down_8 = limma_result %>%
  filter(time == 8 & regulation == "Down") %>%
  ggplot(aes(x=fct_reorder(gene, logFC), y=-logFC, fill=as_factor(sign(logFC)))) +
  geom_col() +
  labs(x=NULL, y="Down-regulated logFC") +
  scale_fill_manual(values = aachen_color("blue")) +
  theme(legend.position = "none",
        axis.text = element_text(size=10),
        axis.title = element_text(size=12)) + 
  scale_x_discrete(position = "top") +
  scale_y_reverse(breaks = c(0,4,8)) +
  # scale_y_reverse(breaks = function(x) unique(floor(pretty(seq(0, (max(x) + 1) * 1.1))))) +
  coord_flip()

up_24 = limma_result %>%
  filter(time == 24 & regulation == "Up") %>%
  ggplot(aes(x=fct_reorder(gene, logFC), y=logFC, fill=as_factor(sign(logFC)))) +
  geom_col() +
  labs(x=NULL, y="Up-regulated logFC") +
  scale_fill_manual(values = aachen_color("green")) +
  scale_y_continuous(position = "right", 
                     breaks = function(x) unique(floor(pretty(seq(0, (max(x) + 1) * 1.1))))) +
  theme(legend.position = "none",
        axis.text = element_text(size=10),
        axis.title = element_text(size=12)) +
  coord_flip()


down_24 = limma_result %>%
  filter(time == 24 & regulation == "Down") %>%
  ggplot(aes(x=fct_reorder(gene, logFC), y=-logFC, fill=as_factor(sign(logFC)))) +
  geom_col() +
  labs(x=NULL, y="Down-regulated logFC") +
  scale_fill_manual(values = aachen_color("blue")) +
  theme(legend.position = "none",
        axis.text = element_text(size=10),
        axis.title = element_text(size=12)) + 
  scale_x_discrete(position = "top") +
  scale_y_reverse(breaks = c(0,4,8)) +
  # scale_y_reverse(breaks = function(x) unique(floor(pretty(seq(0, (max(x) + 1) * 1.1))))) +
  coord_flip()

up_48= limma_result %>%
  filter(time == 48 & regulation == "Up") %>%
  ggplot(aes(x=fct_reorder(gene, logFC), y=logFC, fill=as_factor(sign(logFC)))) +
  geom_col() +
  labs(x=NULL, y="Up-regulated logFC") +
  scale_fill_manual(values = aachen_color("green")) +
  scale_y_continuous(position = "right",
                     breaks = function(x) unique(floor(pretty(seq(0, (max(x) + 1) * 1.1))))) +
  theme(legend.position = "none",
        axis.text = element_text(size=10),
        axis.title = element_text(size=12)) +
  coord_flip()


down_48 = limma_result %>%
  filter(time == 48 & regulation == "Down") %>%
  ggplot(aes(x=fct_reorder(gene, logFC), y=-logFC, fill=as_factor(sign(logFC)))) +
  geom_col() +
  labs(x=NULL, y="Down-regulated logFC") +
  scale_fill_manual(values = aachen_color("blue")) +
  theme(legend.position = "none",
        axis.text = element_text(size=10),
        axis.title = element_text(size=12)) + 
  scale_x_discrete(position = "top") +
  scale_y_reverse(breaks = function(x) unique(floor(pretty(seq(0, (max(x) + 1) * 1.1))))) +
  coord_flip()

saveRDS(list(up_8 = up_8, down_8 = down_8, 
             up_24 = up_24, down_24 = down_24,
             up_48 = up_48, down_48 = down_48),
             "fig/ccl4_experiment/top_ccl4_degs.rds")
```

##### Top 40 Oil genes
```{r "top 40 oil genes"}
limma_result = readRDS("output/ccl4_experiment/limma_result.rds") %>%
  filter(contrast_reference == "oil") %>%
  mutate(time = str_extract(contrast, "\\d*$"),
         time = ordered(as.integer(time))) %>%
  mutate(label = fct_inorder(str_c("Week", time, sep=" "))) %>%
  mutate(regulation = str_to_title(regulation)) %>%
  mutate(regulation = factor(regulation, levels = c("Up", "Down", "No"))) %>%
  filter(regulation != "No") %>%
  group_by(time, regulation) %>%
  top_n(20, abs(logFC)) %>%
  ungroup()

up_8 = limma_result %>%
  filter(time == 8 & regulation == "Up") %>%
  ggplot(aes(x=fct_reorder(gene, logFC), y=logFC, fill=as_factor(sign(logFC)))) +
  geom_col() +
  labs(x=NULL, y="Up-regulated logFC") +
  scale_fill_manual(values = aachen_color("green")) +
  scale_y_continuous(position = "right",
                     breaks = function(x) unique(floor(pretty(seq(0, (max(x) + 1) * 1.1))))) +
  theme(legend.position = "none",
        axis.text = element_text(size=10),
        axis.title = element_text(size=12)) +
  coord_flip()


down_8 = limma_result %>%
  filter(time == 8 & regulation == "Down") %>%
  ggplot(aes(x=fct_reorder(gene, logFC), y=-logFC, fill=as_factor(sign(logFC)))) +
  geom_col() +
  labs(x=NULL, y="Down-regulated logFC") +
  scale_fill_manual(values = aachen_color("blue")) +
  theme(legend.position = "none",
        axis.text = element_text(size=10),
        axis.title = element_text(size=12)) + 
  scale_x_discrete(position = "top") +
  scale_y_reverse(breaks = function(x) unique(floor(pretty(seq(0, (max(x) + 1) * 1.1))))) +
  coord_flip()

up_48= limma_result %>%
  filter(time == 48 & regulation == "Up") %>%
  ggplot(aes(x=fct_reorder(gene, logFC), y=logFC, fill=as_factor(sign(logFC)))) +
  geom_col() +
  labs(x=NULL, y="Up-regulated logFC") +
  scale_fill_manual(values = aachen_color("green")) +
  scale_y_continuous(position = "right",
                     breaks = function(x) unique(floor(pretty(seq(0, (max(x) + 1) * 1.1))))) +
  theme(legend.position = "none",
        axis.text = element_text(size=10),
        axis.title = element_text(size=12)) +
  coord_flip()


down_48 = limma_result %>%
  filter(time == 48 & regulation == "Down") %>%
  ggplot(aes(x=fct_reorder(gene, logFC), y=-logFC, fill=as_factor(sign(logFC)))) +
  geom_col() +
  labs(x=NULL, y="Down-regulated logFC") +
  scale_fill_manual(values = aachen_color("blue")) +
  theme(legend.position = "none",
        axis.text = element_text(size=10),
        axis.title = element_text(size=12)) + 
  scale_x_discrete(position = "top") +
  scale_y_reverse(breaks = function(x) unique(floor(pretty(seq(0, (max(x) + 1) * 1.1))))) +
  coord_flip()

saveRDS(list(up_8 = up_8, down_8 = down_8, 
             up_48 = up_48, down_48 = down_48),
             "fig/ccl4_experiment/top_oil_degs.rds")

```


##### CCL4 volcano plot
```{r "ccl4 volcano plot"}
limma_result = readRDS("output/ccl4_experiment/limma_result.rds") %>%
  filter(contrast_reference == "ccl4") %>%
  mutate(time = str_extract(contrast, "\\d*$"),
         time = ordered(as.integer(time)/4)) %>%
  mutate(label = fct_inorder(str_c("Month", time, sep=" "))) %>%
  mutate(regulation = str_to_title(regulation)) %>%
  mutate(regulation = factor(regulation, levels = c("Up", "Down", "No")))

deg_count = limma_result %>%
  filter(regulation != "No") %>%
  add_count(time, label, regulation) %>%
  mutate(logFC = case_when(regulation == "Up" ~ 0.75*max(logFC),
                          regulation == "Down" ~ 0.75*min(logFC)),
         pval = 0.4) %>%
  distinct(label, n, logFC, pval, regulation)

# ccl4
volcano_plot_ccl4 = limma_result %>%
  ggplot(aes(x=logFC, y=-log10(pval), color=regulation, alpha = regulation)) +
  geom_point(size=2) +
  facet_rep_wrap(~label) +
  my_theme(grid = "y") +
  scale_color_manual(values = aachen_color(c("green", "blue", "black50"))) +
  scale_alpha_manual(values = c(0.7,0.7,0.2), guide ="none") +
  labs(y = expression(-log['10']*"(p-value)"), color = "Regulation") +
  geom_text(data = deg_count, aes(label = n), size=14/(14/5), show.legend = F) +
  theme(legend.position = "top")

saveRDS(volcano_plot_ccl4, "fig/ccl4_experiment/volcano_plot_ccl4.rds")
```

##### Oil volcano plot
```{r "oil volcano plot"}
limma_result = readRDS("output/ccl4_experiment/limma_result.rds") %>%
  filter(contrast_reference == "oil") %>%
  mutate(time = str_extract(contrast, "\\d*$"),
         time = ordered(as.integer(time)/4)) %>%
  mutate(label = fct_inorder(str_c("Month", time, sep=" "))) %>% 
  mutate(regulation = str_to_title(regulation)) %>%
  mutate(regulation = factor(regulation, levels = c("Up", "Down", "No")))

deg_count = limma_result %>%
  filter(regulation != "No") %>%
  add_count(time, label, regulation) %>%
  mutate(logFC = case_when(regulation == "Up" ~ 0.75*max(logFC),
                          regulation == "Down" ~ 0.75*min(logFC)),
         pval = 0.4) %>%
  distinct(label, n, logFC, pval, regulation)

volcano_plot_oil = limma_result %>%
  ggplot(aes(x=logFC, y=-log10(pval), color=regulation, alpha = regulation)) +
  geom_point(size=2) +
  facet_rep_wrap(~label, repeat.tick.labels = F) +
  my_theme(grid = "y") +
  scale_color_manual(values = aachen_color(c("green", "blue", "black50"))) +
  scale_alpha_manual(values = c(0.7,0.7,0.2), guide ="none") +
  labs(y = expression(-log['10']*"(p-value)"), color = "Regulation") +
  geom_text(data = deg_count, aes(label = n), size=14/(14/5), show.legend = F) +
  theme(legend.position = "top")

saveRDS(volcano_plot_oil, "fig/ccl4_experiment/volcano_plot_oil.rds")
```

##### PROGENy heatmap CCL4
```{r "ccl4 - progeny}
progeny_scores = readRDS("output/ccl4_experiment/progeny_scores_contrast_wise.rds") %>%
  filter(contrast_reference == "ccl4") %>%
  transmute(pathway, activity, time = str_c("Month", time/4, sep = " ")) %>%
  mutate(time = factor(time, levels = c("Month 2", "Month 6", "Month 12")))

mat = progeny_scores %>%
  spread(time, activity) %>%
  data.frame(row.names = 1, check.names = F, stringsAsFactors = F)

color_breaks = make_div_color_breaks(mat, 
                                     aachen_color("blue"), 
                                     "white", 
                                     aachen_color("green"))

progeny_ccl4 = pheatmap(mat,fontsize=12, fontsize_row = 12, 
                        color=color_breaks$colors, breaks = color_breaks$breaks, 
                        angle_col = 45, cluster_cols = F, border_color = NA,
                        cellwidth = 25) %>%
  as.ggplot()

saveRDS(progeny_ccl4, "fig/ccl4_experiment/progeny_ccl4.rds")
```

##### DoRothEA heatmap CCL4
```{r "ccl4 - dorothea}
dorothea_scores = readRDS(
  "output/ccl4_experiment/dorothea_scores_contrast_wise.rds"
  ) %>%
  filter(contrast_reference == "ccl4") %>%
  transmute(tf, confidence, activity, 
            time = str_c("Month", time/4, sep = " ")) %>%
  mutate(time = factor(time, levels = c("Month 2", "Month 6", "Month 12")))

top_tfs = dorothea_scores %>%
  filter(confidence %in% c("A", "B")) %>%
  filter(time == "Week 48") %>%
  group_by(sign(activity)) %>%
  top_n(10, abs(activity)) %>%
  ungroup() %>%
  distinct(tf)

literature_tfs = tibble(tf = c("Hnf1a","Foxo3","Nr2f2","Esr2","Stat5b",
                               "Stat1", "Egr1", "Atf1", "Hif1a", "Trp53"))

interesting_tfs = bind_rows(top_tfs, literature_tfs)

mat = dorothea_scores %>%
  semi_join(interesting_tfs, by="tf") %>%
  select(-confidence) %>%
  spread(time, activity) %>%
  data.frame(row.names = 1, check.names = F, stringsAsFactors = F)

color_breaks = make_div_color_breaks(mat, aachen_color("blue"), "white", aachen_color("green"))

dorothea_ccl4 = pheatmap(mat,fontsize=12, fontsize_row = 12, 
                        color=color_breaks$colors, breaks = color_breaks$breaks, 
                        angle_col = 45, cluster_cols = F, border_color = NA,
                        cellwidth = 25) %>%
  as.ggplot()

saveRDS(dorothea_ccl4, "fig/ccl4_experiment/dorothea_ccl4.rds")
```

##### GO heatmap CCL4
```{r}
gsea_res = readRDS("output/ccl4_experiment/go_enrichment.rds") %>%
  filter(contrast_reference == "ccl4") %>%
  transmute(geneset, category_subcode, pval, padj, NES, 
            time = str_c("Month", as.numeric(as.character(time))/4, sep = " ")) %>%
  mutate(time = factor(time, levels = c("Month 2", "Month 6", "Month 12")))

interesting_go_terms = tibble(
  geneset = c("GO_ACTIN_BINDING","GO_ANGIOGENESIS", "GO_BIOLOGICAL_ADHESION", 
              "GO_CALCIUM_ION_TRANSPORT", "GO_CELL_CYCLE", "GO_CELL_DEATH", 
              "GO_CELL_DEVELOPMENT", "GO_CELL_MOTILITY", "GO_ENDOCYTOSIS", 
              "GO_IMMUNE_RESPONSE", "GO_INTEGRIN_BINDING", 
              "GO_LEUKOCYTE_ACTIVATION", 
              "GO_MICROTUBULE_CYTOSKELETON_ORGANIZATION", 
              "GO_POSITIVE_REGULATION_OF_CELL_DEATH", 
              "GO_POSITIVE_REGULATION_OF_MAPK_CASCADE")
)

df = gsea_res %>%
  semi_join(interesting_go_terms) %>%
  mutate(geneset = str_remove(geneset, "GO_")) %>%
  mutate(geneset = str_replace_all(geneset, "_", " ")) %>%
  mutate(geneset = str_to_title(geneset)) %>%
  select(-c(pval, padj))

anno = df %>%
  distinct(geneset, category_subcode) %>%
  data.frame(row.names = 1, check.names = F, stringsAsFactors = F)

mat = df %>%
  select(-category_subcode) %>%
  spread(time, NES) %>%
  data.frame(row.names = 1, check.names = F, stringsAsFactors = F)


color_breaks = make_div_color_breaks(mat, 
                                     aachen_color("blue"), 
                                     "white", 
                                     aachen_color("green"))

go_ccl4 = pheatmap(mat,fontsize=12, fontsize_row = 12, 
                        color=color_breaks$colors, breaks = color_breaks$breaks, 
                        angle_col = 45, cluster_cols = F, border_color = NA,
                        cellwidth = 25
                   ) %>%
  as.ggplot()

saveRDS(go_ccl4, "fig/ccl4_experiment/go_ccl4.rds")
```


### Zonation Analysis
#### Overview of individual gene sets
```{r}
gene_sets = list.files("output/zonation/individual_gene_sets", 
                       pattern = "gene_sets", full.names = T) %>%
  map_df(readRDS) %>%
  filter(source != "halpern2018")

geneset_overview = gene_sets %>%
  count(source, zonation) %>%
  mutate_if(is.character, str_to_title) %>%
  ggplot(aes(x=source, y=n, fill=zonation)) +
  geom_bar(stat = "identity", position=position_dodge()) +
  labs(x = "Literature source", y="Count", fill="Zonation") +
  scale_fill_manual(values = aachen_color(c("violet", "turquoise"))) +
  my_theme("y") +
  theme(legend.position = "top")

saveRDS(geneset_overview, "fig/zonation/geneset_overview.rds")
```

#### Overlap of individual gene set
```{r}
gene_sets = list.files("output/zonation/individual_gene_sets", 
                       pattern = "gene_sets", full.names = T) %>%
  map_df(readRDS) %>%
  filter(source != "halpern2018")

# periportal
periportal = gene_sets %>% 
  filter(zonation == "periportal") %>%
  select(gene, source)

pp_h = periportal %>% filter(source == "halpern")
pp_b = periportal %>% filter(source == "braeuning")
pp_s = periportal %>% filter(source == "saito")

pp_hb = periportal %>% 
  filter(source %in% c("halpern", "braeuning")) %>% 
  group_split(source, keep = F) %>% 
  reduce(intersect)
pp_bs = periportal %>% 
  filter(source %in% c("braeuning", "saito")) %>%
  group_split(source, keep = F) %>% 
  reduce(intersect)
pp_hs = periportal %>% 
  filter(source %in% c("halpern", "saito")) %>% 
  group_split(source, keep = F) %>% 
  reduce(intersect)

pp_hbs = periportal %>% 
  group_split(source, keep = F) %>% 
  reduce(intersect)

grid.newpage()
pp_venn = draw.triple.venn(
  area1 = nrow(pp_h), area2 = nrow(pp_b), area3 = nrow(pp_s), 
  n12 = nrow(pp_hb), n23 = nrow(pp_bs), n13 = nrow(pp_hs), 
  n123 = nrow(pp_hbs),
  category = c("Halpern","Braeuning", "Saito"),
  lty = "blank",
  cex = 1,
  fontfamily = rep("sans", 7),
  fill = aachen_color(c("purple", "orange", "petrol")),
  cat.col = aachen_color(c("purple", "orange", "petrol")),
  cat.cex = 1.1,
  cat.fontfamily = rep("sans", 3)
) %>%
  as_ggplot() %>%
  grid.arrange(top = textGrob("Periportal", gp=gpar(fontsize=15, 
                                                    fontface = "bold")))

saveRDS(pp_venn, "fig/zonation/pp_venn.rds")

# pericentral
pericentral = gene_sets %>% 
  filter(zonation == "pericentral") %>%
  select(gene, source)

pc_h = pericentral %>% filter(source == "halpern")
pc_b = pericentral %>% filter(source == "braeuning")
pc_s = pericentral %>% filter(source == "saito")

pc_hb = pericentral %>% 
  filter(source %in% c("halpern", "braeuning")) %>% 
  group_split(source, keep = F) %>% 
  reduce(intersect)
pc_bs = pericentral %>% 
  filter(source %in% c("braeuning", "saito")) %>%
  group_split(source, keep = F) %>% 
  reduce(intersect)
pc_hs = pericentral %>% 
  filter(source %in% c("halpern", "saito")) %>% 
  group_split(source, keep = F) %>% 
  reduce(intersect)

pc_hbs = pericentral %>% 
  group_split(source, keep = F) %>% 
  reduce(intersect)

grid.newpage()
pc_venn = draw.triple.venn(
  area1 = nrow(pc_h), area2 = nrow(pc_b), area3 = nrow(pc_s), 
  n12 = nrow(pc_hb), n23 = nrow(pc_bs), n13 = nrow(pc_hs), 
  n123 = nrow(pc_hbs),
  category = c("Halpern","Braeuning", "Saito"),
  fontfamily = rep("sans", 7),
  lty = "blank",
  cex = 1,
  fill = aachen_color(c("purple", "orange", "petrol")),
  cat.col = aachen_color(c("purple", "orange", "petrol")),
  cat.cex = 1.1,
  cat.fontfamily = rep("sans", 3)
) %>% 
  as_ggplot() %>%
  grid.arrange(top = textGrob("Pericentral", gp=gpar(fontsize=15, 
                                                     fontface = "bold")))

saveRDS(pc_venn, "fig/zonation/pc_venn.rds")
```

#### Enrichment of consensus gene sets with CCl4 signature
```{r}
gsea_res_ccl4 = readRDS("output/zonation/consensus_and_ind_gsea_res.rds") %>%
  filter(contrast_reference == "ccl4",
         source == "consensus") %>%
  mutate(time = ordered(as.numeric(as.character(time))/4))
         

consensus_gsea_ccl4_bar = gsea_res_ccl4 %>% 
  mutate(zonation = str_to_title(zonation)) %>%
  ggplot(aes(x=time, y=NES, fill=zonation)) +
  geom_col() +
  # coord_flip() +
  scale_fill_manual(values = aachen_color(c("violet", "turquoise"))) +
  labs(x = "Time in months", fill = "Zonation") +
  theme(legend.position = "top") +
  geom_text(aes(label = str_c("p = ", scientific(pval,3), sep = " "), 
                y=0.9*NES), color="white", position=position_dodge(width=0), 
            size=4.5) +
  my_theme(grid = "y")

saveRDS(consensus_gsea_ccl4_bar, "fig/zonation/consensus_pc_pp_gsea_ccl4.rds")
```

#### GSEA plots CCL4
```{r}
gsea_res_ccl4 = readRDS("output/zonation/consensus_and_ind_gsea_res.rds") %>%
  filter(contrast_reference == "ccl4",
         source == "consensus") %>%
  mutate(time = ordered(as.numeric(as.character(time))/4))

zonation_gene_sets = readRDS("output/zonation/consensus_gene_sets.rds") %>%
  group_by(zonation) %>%
  summarise(gene_set = list(gene)) %>%
  deframe()

gsea_plots_ccl4 = gsea_res_ccl4 %>% 
  inner_join(enframe(zonation_gene_sets, "zonation", "geneset"), 
             by="zonation") %>%
  mutate(gsea_plots = pmap(., .f = make_gsea_plot))

saveRDS(gsea_plots_ccl4, "fig/zonation/gsea_plots_ccl4.rds")
```

#### Leading edge analysis
```{r}
df = readRDS("output/zonation/consensus_and_ind_gsea_res.rds") %>%
  filter(contrast_reference == "ccl4",
         source == "consensus") %>%
  unnest(leadingEdge) %>%
  distinct(time, zonation, gene = leadingEdge) %>%
  group_by(time, zonation) %>%
  mutate(importance = row_number()) %>%
  ungroup() %>%
  add_count(gene, zonation) %>%
  filter(n>=2) %>%
  distinct(zonation, gene, importance) %>%
  group_by(zonation, gene) %>% 
  summarise(m = mean(importance)) %>%
  group_by(zonation) %>%
  top_n(15, -m) %>%
  ungroup() %>%
  select(-m)

limma_result = readRDS("output/ccl4_experiment/limma_result.rds") %>%
  filter(contrast_reference == "ccl4") %>%
  mutate(time = str_extract(contrast, "\\d*$"),
         time = ordered(as.integer(time)/4)) %>%
  mutate(label = fct_inorder(str_c("Month", time, sep=" "))) %>%
  inner_join(df, by="gene") %>%
  select(gene, logFC, label, zonation) %>%
  spread(label, logFC)


mat = limma_result %>%
  select(-zonation) %>%
  data.frame(row.names = 1, check.names = F, stringsAsFactors = F)

anno = limma_result %>%
  select(gene, Zonation = zonation) %>%
  mutate(Zonation = str_to_title(Zonation)) %>%
  data.frame(row.names=1, check.names = F, stringsAsFactors = F)


color_breaks = make_div_color_breaks(mat, aachen_color("blue"), "white", aachen_color("green"))

leading_edge_hmap = pheatmap(
  mat,fontsize=12, fontsize_row = 12, color=color_breaks$colors, 
  breaks = color_breaks$breaks, angle_col = 45, cluster_cols = F, 
  border_color = NA, cellwidth = 25, annotation_row = anno, 
  annotation_colors = list(Zonation = c(Pericentral = aachen_color("violet"),
                                        Periportal = aachen_color("turquoise")))) %>%
  as.ggplot()

saveRDS(leading_edge_hmap, "fig/zonation/leading_edge_hmap.rds")
```

#### GO terms table of overlap
```{r}
df = readRDS("output/zonation/ora_result_of_overlap.rds") %>%
  filter(contrast_reference == "ccl4")

interesting_genesets = tibble(geneset = c(
  "Nr0b2",
  "Nr1h4",
  "Tbx3",
  "Atf1",
  "GO_CARBOXYLIC_ESTER_HYDROLASE_ACTIVITY",
  "GO_MONOCARBOXYLIC_ACID_METABOLIC_PROCESS",
  "GO_GLUTAMINE_FAMILY_AMINO_ACID_CATABOLIC_PROCESS",
  "GO_LONG_CHAIN_FATTY_ACID_METABOLIC_PROCESS",
  "GO_GLUTAMATE_METABOLIC_PROCESS",
  "GO_EPOXYGENASE_P450_PATHWAY",
  "GO_PHOSPHOLIPID_TRANSPORT",
  "GO_LIPID_METABOLIC_PROCESS",
  "GO_POSITIVE_REGULATION_OF_TRIGLYCERIDE_LIPASE_ACTIVITY",
  "KEGG_ARGININE_AND_PROLINE_METABOLISM",
  "KEGG_PRIMARY_BILE_ACID_BIOSYNTHESIS"
))

t = df %>%
  filter(regulation == "down" & zonation == "pericentral" | 
           regulation == "up" & zonation == "periportal") %>%
  semi_join(interesting_genesets, by="geneset") %>%
  select(regulation, zonation, category_subcode, geneset, p.value) %>%
  arrange(geneset) %>%
  filter(p.value <= 0.02) %>%
  filter(!(geneset == "KEGG_PRIMARY_BILE_ACID_BIOSYNTHESIS" & 
             zonation == "periportal")) %>%
  arrange(zonation, regulation, p.value) %>%
  mutate(geneset = str_replace(geneset, "GO_", ""),
         geneset = str_replace(geneset, "KEGG_", ""),
         geneset = str_replace_all(geneset, "_", " "),
         geneset = str_to_title(geneset),
         regulation = str_to_title(regulation),
         zonation = str_to_title(zonation)) %>%
mutate(
  category_subcode = case_when(
           category_subcode == "mf" ~ "GO molecular function",
           category_subcode == "bp" ~ "GO biological process",
           category_subcode == "dorothea" ~ "DoRothEA TF")
) %>%
  mutate(p.value = scientific(p.value)) %>%
  rename(Regulation = regulation, Zonation = zonation, 
         Category = category_subcode, "Gene set" = geneset, p = p.value) %>%
  select(-Category)

t1 = ttheme_default(
  core=list(
    bg_params = list(fill=c("white"))#,
    # fg_params = list(col=c("orange", "blue"))
    ),
  colhead = list(
    bg_params = list(fill = c("white"))
    )
)
grid.newpage()
tab = t %>%
  gridExtra::tableGrob(rows = NULL, theme = t1) %>%
  gtable::gtable_add_grob(
    .,
    grobs = grid::segmentsGrob( # line across the bottom
      x0 = unit(0,"npc"),
      y0 = unit(0,"npc"),
      x1 = unit(1,"npc"),
      y1 = unit(0,"npc"),
      gp = gpar(lwd = 3.0)),
    t = 1, b = 1, l = 1, r = ncol(.)) %>%
  gtable::gtable_add_grob(
    .,
    grobs = grid::segmentsGrob( # line across the bottom
      x0 = unit(0,"npc"),
      y0 = unit(0,"npc"),
      x1 = unit(1,"npc"),
      y1 = unit(0,"npc"),
      gp = gpar(lwd = 3.0)),
    t = nrow(.), b = 1, l = 1, r = ncol(.))


# color the words pericentral and periportal
# tab$grobs[24][[1]][["gp"]] = gpar(col = aachen_color("violet"))
tab$grobs[25][[1]][["gp"]] = gpar(col = aachen_color("violet"))
tab$grobs[26][[1]][["gp"]] = gpar(col = aachen_color("violet"))
tab$grobs[27][[1]][["gp"]] = gpar(col = aachen_color("violet"))
tab$grobs[28][[1]][["gp"]] = gpar(col = aachen_color("violet"))
tab$grobs[29][[1]][["gp"]] = gpar(col = aachen_color("violet"))
tab$grobs[30][[1]][["gp"]] = gpar(col = aachen_color("violet"))
tab$grobs[31][[1]][["gp"]] = gpar(col = aachen_color("violet"))
tab$grobs[32][[1]][["gp"]] = gpar(col = aachen_color("violet"))
tab$grobs[33][[1]][["gp"]] = gpar(col = aachen_color("violet"))
tab$grobs[34][[1]][["gp"]] = gpar(col = aachen_color("violet"))
tab$grobs[35][[1]][["gp"]] = gpar(col = aachen_color("violet"))
tab$grobs[36][[1]][["gp"]] = gpar(col = aachen_color("turquoise"))
tab$grobs[37][[1]][["gp"]] = gpar(col = aachen_color("turquoise"))
tab$grobs[38][[1]][["gp"]] = gpar(col = aachen_color("turquoise"))
tab$grobs[39][[1]][["gp"]] = gpar(col = aachen_color("turquoise"))
tab$grobs[40][[1]][["gp"]] = gpar(col = aachen_color("turquoise"))

# make p italic
tab$grobs[4][[1]][["gp"]] = gpar(fontface = 4L)

# color KEGG pathway
tab$grobs[48][[1]][["gp"]] = gpar(col = aachen_color("maygreen"))
tab$grobs[50][[1]][["gp"]] = gpar(col = aachen_color("maygreen"))

# color mf
tab$grobs[43][[1]][["gp"]] = gpar(col = aachen_color("orange"))

# color bp
tab$grobs[41][[1]][["gp"]] = gpar(col = aachen_color("red"))
tab$grobs[44][[1]][["gp"]] = gpar(col = aachen_color("red"))
tab$grobs[45][[1]][["gp"]] = gpar(col = aachen_color("red"))
tab$grobs[46][[1]][["gp"]] = gpar(col = aachen_color("red"))
tab$grobs[47][[1]][["gp"]] = gpar(col = aachen_color("red"))
tab$grobs[49][[1]][["gp"]] = gpar(col = aachen_color("red"))
tab$grobs[52][[1]][["gp"]] = gpar(col = aachen_color("red"))
tab$grobs[53][[1]][["gp"]] = gpar(col = aachen_color("red"))
tab$grobs[54][[1]][["gp"]] = gpar(col = aachen_color("red"))

# color tf
tab$grobs[42][[1]][["gp"]] = gpar(col = aachen_color("purple"))
tab$grobs[51][[1]][["gp"]] = gpar(col = aachen_color("purple"))
tab$grobs[55][[1]][["gp"]] = gpar(col = aachen_color("purple"))
tab$grobs[56][[1]][["gp"]] = gpar(col = aachen_color("purple"))

# make legend
legend_plot = tibble(source = factor(c("Molecular function (GO)", 
                                       "Biological process (GO)", 
                                       "TF (DoRothEA)", 
                                       "Pathway (KEGG)"), 
                                     levels = c("Biological process (GO)", 
                                                "Molecular function (GO)", 
                                                "Pathway (KEGG)", 
                                                "TF (DoRothEA)"))) %>%
  ggplot(aes(x=1, y=1,color=source, fill=source, label = source)) +
  geom_point(size=3, shape=15) +
  # geom_text(size=4) +
  scale_color_manual(values = aachen_color(c("red", "orange", "maygreen", 
                                           "purple"))) +
  theme(legend.title = element_blank(),
        legend.position = "bottom",
        legend.margin=margin(c(t=-50,r=0,b=0,l=0))) +
  guides(color = guide_legend(ncol=2))


tab_with_legend = plot_grid(
  tab, 
  get_legend(legend_plot),
  ncol=1, rel_heights = c(1,0.1))
  
saveRDS(tab_with_legend, "fig/zonation/tab.rds")
plot(tab)

ind <- find_cell(tab, 9, 3, "core-fg")


find_cell <- function(table, row, col, name="core-fg"){
  l <- table$layout
  which(l$t==row & l$l==col & l$name==name)
}


```


#### Enrichment of individual gene sets with ccl4 signature
```{r}
gsea_res_ccl4_ind = readRDS(
  "output/zonation/consensus_and_ind_gsea_res.rds"
  ) %>%
  filter(contrast_reference == "ccl4",
         source != "consensus") %>%
  mutate(time = ordered(as.numeric(as.character(time))/4))

ind_gsea_ccl4_bar = gsea_res_ccl4_ind %>% 
  mutate(zonation = str_to_title(zonation),
         source = str_to_title(source)) %>%
  ggplot(aes(x=time, y=NES, fill=zonation)) +
  geom_col(position=position_dodge()) +
  facet_rep_wrap(~source) +
  scale_fill_manual(values = aachen_color(c("violet", "turquoise"))) +
  labs(x = "Time in months", fill = "Zonation") +
  theme(legend.position = "top") +
  geom_text(aes(label = gtools::stars.pval(pval), y=0.9*NES), color="white", 
            position=position_dodge(width=0.9), size=5) +
  my_theme("y")

saveRDS(ind_gsea_ccl4_bar, "fig/zonation/ind_pc_pp_gsea_ccl4.rds")
```


