---
title: "Script to plot individual figures"
author: "Christian Holland"
date: "6/23/2019"
output: html_document
---

```{r "knitr config", cache=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::knit(..., quiet = TRUE)
```
### Libraries and sources
These libraries and sources are used in this analysis 
```{r "setup", message=F}
library(tidyverse)
library(cowplot)
library(tidylog)
library(RWTHColorPalette) # devtools::install_github("christianholland/RWTHColorPalette")
library(ggpubr)
library(UpSetR)
library(VennDiagram)
library(lemon)
library(gridExtra)
library(lemon)
library(scales)
library(ggplotify)
library(pheatmap)
library(fgsea)


options("tidylog.display" = list(print))
theme_set(theme_cowplot())
source("src/my_ggplot_themes.R")

```
### Utility functions
```{r}
make_gsea_plot = function(data, geneset, NES, padj, time, zonation, ...) {
    stats_df = data %>%
      rename(stat = statistic) %>%
      arrange(-stat) %>% 
      mutate(stat = stat/max(abs(stat)),
             rank = row_number()) 
    
    rnk = stats_df$rank
    
    adj_stats = stats_df %>%
      select(gene, stat) %>%
      deframe()
  
    pathway = match(geneset, names(adj_stats)) %>%
      na.omit() %>%
      as.vector() %>%
      unname() %>%
      sort()
    
    
    gseaRes = calcGseaStat(adj_stats, selectedStats = pathway, returnAllExtremes = TRUE)
    bottoms = gseaRes$bottoms
    tops = gseaRes$tops
    n = length(adj_stats)
    xs = as.vector(rbind(pathway - 1, pathway))
    ys = as.vector(rbind(bottoms, tops))
    
    enrichment = tibble(rank = c(0, xs, n + 1), 
                        running_sum = c(0, ys, 0))
    ticks = tibble(rank = pathway, 
                   max_rank = max(rnk))
    color_df = enframe(rnk) %>%
      mutate(bin = factor(ntile(value, 10)))
    
    tmp = enrichment %>% 
      filter(abs(running_sum) == max(abs(running_sum)))
    
    
    g1 = ggplot(data = enrichment, aes(x = rank, y = running_sum)) + 
      geom_hline(yintercept = max(tops), colour = rwth_color("bordeaux"), linetype = "dashed") +
      geom_hline(yintercept = min(bottoms), colour =rwth_color("bordeaux"), linetype = "dashed") +
      geom_hline(yintercept = 0, colour = "black") +
      geom_segment(aes(x = tmp$rank, y = 0, xend = tmp$rank, yend = tmp$running_sum), 
                   linetype = "dashed") +
      geom_line(size=1, color = if_else(zonation == "periportal", rwth_color("turquoise"), rwth_color("violet"))) +
      lims(x = c(0, max(rnk) + 1)) +
      labs(y = "Enrichment score") +
      background_grid(major = "y", minor = "none", size.major = 0.4) +
      theme(axis.line.x = element_blank(),
            axis.title.x = element_blank(),
            axis.ticks.x = element_blank(),
            axis.text.x = element_blank(),
            plot.margin=unit(c(1,1,-0.25,1), "cm"),
            legend.position = "none",
            title = element_text(size=16),
            axis.text = element_text(size=14),
            plot.title = element_text(hjust = 0.5)
            #axis.line.y = element_blank()
            ) +
      annotate("text", x = c(1,6000,12000) , y = c(max(tops)+0.05, max(tops)+0.05, max(tops)+0.05), label = c(
        str_c("NES:", round(NES,2), sep = " "),
        str_c("FDR: ", scientific(padj,3), sep = " "),
        str_c("n:", length(pathway), sep=" ")),
        size=5, hjust = 0) +
      ggtitle(str_c("Week:",time, "-", str_to_title(zonation), sep = " "))
    
    
    g2 = ggplot(color_df, aes(x=value, y=1)) +
      geom_tile(aes(color=value)) +
      geom_segment(ticks, mapping = aes(x = rank, y = 1.51, xend = rank, yend = 3.51), size = 0.5, color=rwth_color("black"), alpha=0.5) +
      scale_color_gradientn(colours = rwth_color(c("bordeaux", "orange", "yellow", "green", "blue", "violet"))) +
      scale_x_continuous(position = "bottom",
                         limits = c(0, max(rnk) +1),
                         breaks = c(1,2000,4000,6000,8000,
                                    10000,12000, 14000)
                         ) +
      theme(
          # axis.line.y = element_blank(),
          axis.ticks.y = element_blank(),
          axis.text.y = element_text(color="white"),
          axis.text.x = element_text(),
          plot.margin=unit(c(-0.25,1,1,1), "cm"),
          legend.position = "none",
          # axis.ticks.x = element_blank(),
          axis.line.x = element_blank(),
          title = element_text(size=16),
          axis.text = element_text(size=14)
            ) +
      labs(x = "Rank", y=NULL) 
    
    plot_grid(g1, g2, ncol = 1, align = "v", axis="l", rel_heights = c(2,1))
}


make_div_color_breaks = function(mat, low, mid, high) {
  paletteLength = 100
  colors = colorRampPalette(c(low, mid, high))(paletteLength)
  breaks = c(seq(min(mat), 0, length.out=ceiling(paletteLength/2) + 1),
             seq(max(mat)/paletteLength, max(mat), length.out=floor(paletteLength/2)))
  
  return(list(colors = colors, breaks = breaks))
}

```


### CCL4 experiment
#### Exploratory analysis
##### PCA
```{r}
pca = readRDS("output/ccl4_experiment/pca.rds")
explained = pca$explained
coords = pca$coords %>%
  mutate(dose = case_when(dose == "D0.6" ~ "0.6 ml/kg CCL4",
                          dose == "oil" ~ "Oil",
                          dose == "D0" ~ "Untreated")) %>%
  mutate(dose = factor(dose, levels = c("Untreated", "Oil", "0.6 ml/kg CCL4")))

my_palette = colorRampPalette(rwth_color(c("blue", "blue25")))

pca_all_1_2 = ggplot(coords, aes(x=PC1, y=PC2, color=time, shape=dose)) +
  geom_point(size=4) +
  labs(x = str_c("PC1 (", explained[1], "%)"),
       y = str_c("PC2 (", explained[2], "%)"),
       color = "Time in weeks",
       shape = "Treatment") +
  my_theme(grid = "no") +
  # scale_color_manual(values = rwth_color(rev(c("blue", "blue75", "blue50", "blue25")))) +
  scale_color_manual(values = rev(my_palette(4))) +
  scale_shape_manual(values = c(15,16,17), labels = c(expression("Untreated"), expression(Oil), expression(CCl['4']))) +
  theme(
    legend.position = "top",
    legend.text.align = 0,
    legend.box = "vertial",
    legend.title.align = 0.5,
    legend.box.just = "left")

saveRDS(pca_all_1_2, "fig/ccl4_experiment/pca_all.rds")


pca_all_1_2_text = ggplot(coords, aes(x=PC1, y=PC2, label=time, color=dose)) +
  geom_text(size=5, alpha = 0.8) +
  labs(x = str_c("PC1 (", explained[1], "%)"),
       y = str_c("PC2 (", explained[2], "%)"),
       color = "Treatment") +
  my_theme(grid = "no") +
  scale_color_manual(values = rwth_color(c("green", "red", "blue")), labels = c(expression("Untreated"), expression(Oil), expression(CCl['4']))) +
  theme(
    legend.position = "top",
    legend.text.align = 0,
    legend.box = "vertial",
    legend.title.align = 0.5,
    legend.box.just = "left")

saveRDS(pca_all_1_2_text, "fig/ccl4_experiment/pca_all_text.rds")

pca_time_1_2 = ggplot(coords, aes(x=PC1, y=PC2, color=time)) +
  geom_point(size=4) +
  labs(x = str_c("PC1 (", explained[1], "%)"),
       y = str_c("PC2 (", explained[2], "%)"),
       color = "Time in weeks") +
  theme(legend.position = "top") +
  my_theme(grid = "no") +
  scale_color_manual(values = rwth_color(c("bordeaux", "orange", "green", "petrol")))

saveRDS(pca_time_1_2, "fig/ccl4_experiment/pca_time.rds")

pca_dose_1_2 = ggplot(coords, aes(x=PC1, y=PC2, color=dose)) +
  geom_point(size=4) +
  labs(x = paste0("PC1 (", explained[1], "%)"),
       y = paste0("PC2 (", explained[2], "%)"),
       color = "Treatment") +
  theme(legend.position = "top") +
  my_theme(grid = "no") +
  scale_color_manual(values = rwth_color(c("violet", "turquoise", "red")), labels = c(expression("Untreated"), expression(Oil), expression(CCl['4']))) +
  theme(legend.text = element_text(vjust = 0.5))

saveRDS(pca_dose_1_2, "fig/ccl4_experiment/pca_dose.rds")
```

#### Differential gene expression analysis
##### Overview
```{r}
limma_result = readRDS("output/ccl4_experiment/limma_result.rds") %>%
  filter(contrast_reference == "ccl4") %>%
  mutate(time = str_extract(contrast, "\\d*$"),
         time = ordered(as.integer(time))) %>%
  mutate(regulation = str_to_title(regulation)) %>%
  mutate(regulation = factor(regulation, levels = c("Up", "Down", "No"))) %>%
  mutate(time = fct_rev(time))

# stacked barplot showing proportion of regulated genes
deg_overview = limma_result %>% 
  count(time, regulation) %>%
  ggplot(aes(x = time, y=n, fill=regulation)) +
  geom_bar( stat="identity", position="fill") +
  scale_y_continuous(labels = scales::percent) +
  coord_flip(ylim = c(0.9,1)) +
  labs(x= "Time in weeks", y = NULL, fill = "Regulation") +
  my_theme(grid = "x") +
  scale_fill_manual(values = rwth_color(c("green", "red", "black50"))) +
  theme(legend.position = "top")

saveRDS(deg_overview, "fig/ccl4_experiment/deg_overview.rds")
```
##### Top 40 CCl4 genes
```{r}
limma_result = readRDS("output/ccl4_experiment/limma_result.rds") %>%
  filter(contrast_reference == "ccl4") %>%
  mutate(time = str_extract(contrast, "\\d*$"),
         time = ordered(as.integer(time))) %>%
  mutate(label = fct_inorder(str_c("Week", time, sep=" "))) %>%
  mutate(regulation = str_to_title(regulation)) %>%
  mutate(regulation = factor(regulation, levels = c("Up", "Down", "No"))) %>%
  filter(regulation != "No") %>%
  group_by(time, regulation) %>%
  top_n(20, abs(logFC)) %>%
  ungroup()

up_8 = limma_result %>%
  filter(time == 8 & regulation == "Up") %>%
  ggplot(aes(x=fct_reorder(gene, logFC), y=logFC, fill=as_factor(sign(logFC)))) +
  geom_col() +
  labs(x=NULL, y="Up-regulated logFC") +
  scale_fill_manual(values = rwth_color("green")) +
  scale_y_continuous(position = "right", 
                     breaks = function(x) unique(floor(pretty(seq(0, (max(x) + 1) * 1.1))))) +
  theme(legend.position = "none") +
  coord_flip()


down_8 = limma_result %>%
  filter(time == 8 & regulation == "Down") %>%
  ggplot(aes(x=fct_reorder(gene, logFC), y=-logFC, fill=as_factor(sign(logFC)))) +
  geom_col() +
  labs(x=NULL, y="Down-regulated logFC") +
  scale_fill_manual(values = rwth_color("red")) +
  theme(legend.position = "none") + 
  scale_x_discrete(position = "top") +
  scale_y_reverse(breaks = function(x) unique(floor(pretty(seq(0, (max(x) + 1) * 1.1))))) +
  coord_flip()

up_24 = limma_result %>%
  filter(time == 24 & regulation == "Up") %>%
  ggplot(aes(x=fct_reorder(gene, logFC), y=logFC, fill=as_factor(sign(logFC)))) +
  geom_col() +
  labs(x=NULL, y="Up-regulated logFC") +
  scale_fill_manual(values = rwth_color("green")) +
  scale_y_continuous(position = "right", 
                     breaks = function(x) unique(floor(pretty(seq(0, (max(x) + 1) * 1.1))))) +
  theme(legend.position = "none") +
  coord_flip()


down_24 = limma_result %>%
  filter(time == 24 & regulation == "Down") %>%
  ggplot(aes(x=fct_reorder(gene, logFC), y=-logFC, fill=as_factor(sign(logFC)))) +
  geom_col() +
  labs(x=NULL, y="Down-regulated logFC") +
  scale_fill_manual(values = rwth_color("red")) +
  theme(legend.position = "none") + 
  scale_x_discrete(position = "top") +
  scale_y_reverse(breaks = function(x) unique(floor(pretty(seq(0, (max(x) + 1) * 1.1))))) +
  coord_flip()

up_48= limma_result %>%
  filter(time == 48 & regulation == "Up") %>%
  ggplot(aes(x=fct_reorder(gene, logFC), y=logFC, fill=as_factor(sign(logFC)))) +
  geom_col() +
  labs(x=NULL, y="Up-regulated logFC") +
  scale_fill_manual(values = rwth_color("green")) +
  scale_y_continuous(position = "right",
                     breaks = function(x) unique(floor(pretty(seq(0, (max(x) + 1) * 1.1))))) +
  theme(legend.position = "none") +
  coord_flip()


down_48 = limma_result %>%
  filter(time == 48 & regulation == "Down") %>%
  ggplot(aes(x=fct_reorder(gene, logFC), y=-logFC, fill=as_factor(sign(logFC)))) +
  geom_col() +
  labs(x=NULL, y="Down-regulated logFC") +
  scale_fill_manual(values = rwth_color("red")) +
  theme(legend.position = "none") + 
  scale_x_discrete(position = "top") +
  scale_y_reverse(breaks = function(x) unique(floor(pretty(seq(0, (max(x) + 1) * 1.1))))) +
  coord_flip()

saveRDS(list(up_8 = up_8, down_8 = down_8, 
             up_24 = up_24, down_24 = down_24,
             up_48 = up_48, down_48 = down_48),
             "fig/ccl4_experiment/top_ccl4_degs.rds")
```

##### Top 40 Oil genes
```{r}
limma_result = readRDS("output/ccl4_experiment/limma_result.rds") %>%
  filter(contrast_reference == "oil") %>%
  mutate(time = str_extract(contrast, "\\d*$"),
         time = ordered(as.integer(time))) %>%
  mutate(label = fct_inorder(str_c("Week", time, sep=" "))) %>%
  mutate(regulation = str_to_title(regulation)) %>%
  mutate(regulation = factor(regulation, levels = c("Up", "Down", "No"))) %>%
  filter(regulation != "No") %>%
  group_by(time, regulation) %>%
  top_n(20, abs(logFC)) %>%
  ungroup()

up_8 = limma_result %>%
  filter(time == 8 & regulation == "Up") %>%
  ggplot(aes(x=fct_reorder(gene, logFC), y=logFC, fill=as_factor(sign(logFC)))) +
  geom_col() +
  labs(x=NULL, y="Up-regulated logFC") +
  scale_fill_manual(values = rwth_color("green")) +
  scale_y_continuous(position = "right",
                     breaks = function(x) unique(floor(pretty(seq(0, (max(x) + 1) * 1.1))))) +
  theme(legend.position = "none") +
  coord_flip()


down_8 = limma_result %>%
  filter(time == 8 & regulation == "Down") %>%
  ggplot(aes(x=fct_reorder(gene, logFC), y=-logFC, fill=as_factor(sign(logFC)))) +
  geom_col() +
  labs(x=NULL, y="Down-regulated logFC") +
  scale_fill_manual(values = rwth_color("red")) +
  theme(legend.position = "none") + 
  scale_x_discrete(position = "top") +
  scale_y_reverse(breaks = function(x) unique(floor(pretty(seq(0, (max(x) + 1) * 1.1))))) +
  coord_flip()

up_48= limma_result %>%
  filter(time == 48 & regulation == "Up") %>%
  ggplot(aes(x=fct_reorder(gene, logFC), y=logFC, fill=as_factor(sign(logFC)))) +
  geom_col() +
  labs(x=NULL, y="Up-regulated logFC") +
  scale_fill_manual(values = rwth_color("green")) +
  scale_y_continuous(position = "right",
                     breaks = function(x) unique(floor(pretty(seq(0, (max(x) + 1) * 1.1))))) +
  theme(legend.position = "none") +
  coord_flip()


down_48 = limma_result %>%
  filter(time == 48 & regulation == "Down") %>%
  ggplot(aes(x=fct_reorder(gene, logFC), y=-logFC, fill=as_factor(sign(logFC)))) +
  geom_col() +
  labs(x=NULL, y="Down-regulated logFC") +
  scale_fill_manual(values = rwth_color("red")) +
  theme(legend.position = "none") + 
  scale_x_discrete(position = "top") +
  scale_y_reverse(breaks = function(x) unique(floor(pretty(seq(0, (max(x) + 1) * 1.1))))) +
  coord_flip()

saveRDS(list(up_8 = up_8, down_8 = down_8, 
             up_48 = up_48, down_48 = down_48),
             "fig/ccl4_experiment/top_oil_degs.rds")

```

##### Overlap of top 40 CCL4 genes
```{r}
limma_result = readRDS("output/ccl4_experiment/limma_result.rds") %>%
  filter(contrast_reference == "ccl4") %>%
  mutate(time = str_extract(contrast, "\\d*$"),
         time = ordered(as.integer(time))) %>%
  mutate(label = fct_inorder(str_c("Week", time, sep=" "))) %>%
  mutate(regulation = str_to_title(regulation)) %>%
  mutate(regulation = factor(regulation, levels = c("Up", "Down", "No"))) %>%
  filter(regulation != "No")

limma_result %>%
  unite(key, label, regulation, sep="-") %>%
  group_by(key) %>%
  top_n(20, abs(logFC)) %>%
  select(gene, key) %>%
  mutate(val = 1) %>%
  spread(key, val, fill=0) %>%
  data.frame(row.names = 1, check.names = F, stringsAsFactors = F) %>%
  UpSetR::upset(nsets = 6, text.scale = 2, point.size = 4)
```

##### Overlap of top 40 Oil genes
```{r}
limma_result = readRDS("output/ccl4_experiment/limma_result.rds") %>%
  filter(contrast_reference == "oil") %>%
  mutate(time = str_extract(contrast, "\\d*$"),
         time = ordered(as.integer(time))) %>%
  mutate(label = fct_inorder(str_c("Week", time, sep=" "))) %>%
  mutate(regulation = str_to_title(regulation)) %>%
  mutate(regulation = factor(regulation, levels = c("Up", "Down", "No"))) %>%
  filter(regulation != "No") %>%
  group_by(time, regulation) %>%
  top_n(20, abs(logFC)) %>%
  ungroup()

# overlap
limma_result %>%
  unite(label, label, regulation, sep = "-") %>%
  mutate(val = 1) %>%
  select(gene, label, val) %>%
  spread(label, val, fill=0) %>%
  data.frame(row.names = 1, check.names = F, stringsAsFactors = F) %>%
  UpSetR::upset(nsets = 4, text.scale = 2, point.size = 4)

```


##### CCL4 volcano plot (and p-value histogram)
```{r}
limma_result = readRDS("output/ccl4_experiment/limma_result.rds") %>%
  filter(contrast_reference == "ccl4") %>%
  mutate(time = str_extract(contrast, "\\d*$"),
         time = ordered(as.integer(time))) %>%
  mutate(label = fct_inorder(str_c("Week", time, sep=" "))) %>%
  mutate(regulation = str_to_title(regulation)) %>%
  mutate(regulation = factor(regulation, levels = c("Up", "Down", "No")))

deg_count = limma_result %>%
  filter(regulation != "No") %>%
  add_count(time, label, regulation) %>%
  mutate(logFC = case_when(regulation == "Up" ~ 0.75*max(logFC),
                          regulation == "Down" ~ 0.75*min(logFC)),
         pval = 0.4) %>%
  distinct(label, n, logFC, pval, regulation)

# ccl4
volcano_plot_ccl4 = limma_result %>%
  ggplot(aes(x=logFC, y=-log10(pval), color=regulation, alpha = regulation)) +
  geom_point(size=2) +
  facet_rep_wrap(~label) +
  my_theme(grid = "y") +
  scale_color_manual(values = rwth_color(c("green", "red", "black50"))) +
  scale_alpha_manual(values = c(0.7,0.7,0.2), guide ="none") +
  labs(y = expression(-log['10']*"(p-value)"), color = "Regulation") +
  geom_text(data = deg_count, aes(label = n), size=14/(14/5), show.legend = F) +
  theme(legend.position = "top")

saveRDS(volcano_plot_ccl4, "fig/ccl4_experiment/volcano_plot_ccl4.rds")

p_hist_ccl4 = limma_result %>%
  ggplot(aes(x=pval)) +
  geom_histogram(bins = 20, center = 0.025, color="white") +
  facet_grid(~label) +
  labs(x = "p-value", y="Count") +
  my_theme(grid = "y")

saveRDS(p_hist_ccl4, "fig/ccl4_experiment/p_hist_ccl4.rds")
```

##### Oil volcano plot (and p-value histogram)
```{r}
limma_result = readRDS("output/ccl4_experiment/limma_result.rds") %>%
  filter(contrast_reference == "oil") %>%
  mutate(time = str_extract(contrast, "\\d*$"),
         time = ordered(as.integer(time))) %>%
  mutate(label = fct_inorder(str_c("Week", time, sep=" "))) %>% 
  mutate(regulation = str_to_title(regulation)) %>%
  mutate(regulation = factor(regulation, levels = c("Up", "Down", "No")))

deg_count = limma_result %>%
  filter(regulation != "No") %>%
  add_count(time, label, regulation) %>%
  mutate(logFC = case_when(regulation == "Up" ~ 0.75*max(logFC),
                          regulation == "Down" ~ 0.75*min(logFC)),
         pval = 0.4) %>%
  distinct(label, n, logFC, pval, regulation)

volcano_plot_oil = limma_result %>%
  ggplot(aes(x=logFC, y=-log10(pval), color=regulation, alpha = regulation)) +
  geom_point(size=2) +
  facet_rep_wrap(~label, repeat.tick.labels = F) +
  my_theme(grid = "y") +
  scale_color_manual(values = rwth_color(c("green", "red", "black50"))) +
  scale_alpha_manual(values = c(0.7,0.7,0.2), guide ="none") +
  labs(y = expression(-log['10']*"(p-value)"), color = "Regulation") +
  geom_text(data = deg_count, aes(label = n), size=14/(14/5), show.legend = F) +
  theme(legend.position = "top")

saveRDS(volcano_plot_oil, "fig/ccl4_experiment/volcano_plot_oil.rds")

p_hist_oil = limma_result %>%
  ggplot(aes(x=pval)) +
  geom_histogram(bins = 20, center = 0.025, color="white") +
  facet_rep_wrap(~label, repeat.tick.labels = F) +
  labs(x = "p-value", y="Count") +
  my_theme(grid = "y")

saveRDS(p_hist_oil, "fig/ccl4_experiment/p_hist_oil.rds")
```

##### PROGENy heatmap CCL4
```{r}
progeny_scores = readRDS("output/ccl4_experiment/progeny_scores_contrast_wise.rds") %>%
  filter(contrast_reference == "ccl4") %>%
  transmute(pathway, activity, time = str_c("Week", time, sep = " ")) %>%
  mutate(time = factor(time, levels = c("Week 8", "Week 24", "Week 48")))

mat = progeny_scores %>%
  spread(time, activity) %>%
  data.frame(row.names = 1, check.names = F, stringsAsFactors = F)

color_breaks = make_div_color_breaks(mat, rwth_color("blue"), "white", rwth_color("green"))

progeny_ccl4 = pheatmap(mat,fontsize=12, fontsize_row = 12, 
                        color=color_breaks$colors, breaks = color_breaks$breaks, 
                        angle_col = 45, cluster_cols = F, border_color = NA,
                        cellwidth = 25) %>%
  as.ggplot()

saveRDS(progeny_ccl4, "fig/ccl4_experiment/progeny_ccl4.rds")
```

##### PROGENy heatmap Oil
```{r}
progeny_scores = readRDS("output/ccl4_experiment/progeny_scores_contrast_wise.rds") %>%
  filter(contrast_reference == "oil") %>%
  transmute(pathway, activity, time = str_c("Week:", time, sep = " ")) %>%
  mutate(time = factor(time, levels = c("Week: 8", "Week: 24", "Week: 48")))

mat = progeny_scores %>%
  spread(time, activity) %>%
  data.frame(row.names = 1, check.names = F, stringsAsFactors = F)

color_breaks = make_div_color_breaks(mat, rwth_color("blue"), "white", rwth_color("green"))

progeny_oil = pheatmap(mat,fontsize=12, fontsize_row = 12, 
                       color=color_breaks$colors, breaks = color_breaks$breaks, 
                       angle_col = 45, cluster_cols = F, border_color = NA,
                       cellwidth = 25) %>%
  as.ggplot()

saveRDS(progeny_oil, "fig/ccl4_experiment/progeny_oil.rds")
```


##### DoRothEA heatmap CCL4
```{r}
dorothea_scores = readRDS("output/ccl4_experiment/dorothea_scores_contrast_wise.rds") %>%
  filter(contrast_reference == "ccl4") %>%
  transmute(tf, confidence, activity, time = str_c("Week", time, sep = " ")) %>%
  mutate(time = factor(time, levels = c("Week 8", "Week 24", "Week 48")))

mat = dorothea_scores %>%
  group_by(tf) %>%
  mutate(var = var(activity)) %>%
  mutate(mean = mean(activity)) %>%
  mutate(importance = mean) %>%
  ungroup() %>%
  filter(confidence %in% c("A", "B")) %>%
  group_by(direction = sign(importance)) %>%
  top_n(45, abs(importance)) %>% # 3 (for each time point) * 15
  ungroup() %>%
  select(-confidence, -var, -mean, -importance, -direction) %>%
  spread(time, activity) %>%
  data.frame(row.names = 1, check.names = F, stringsAsFactors = F)

color_breaks = make_div_color_breaks(mat, rwth_color("blue"), "white", rwth_color("green"))

dorothea_ccl4 = pheatmap(mat,fontsize=12, fontsize_row = 8, 
                        color=color_breaks$colors, breaks = color_breaks$breaks, 
                        angle_col = 45, cluster_cols = F, border_color = NA,
                        cellwidth = 25) %>%
  as.ggplot()

saveRDS(dorothea_ccl4, "fig/ccl4_experiment/dorothea_ccl4.rds")
```

##### DoRothEA heatmap Oil
```{r}
dorothea_scores = readRDS("output/ccl4_experiment/dorothea_scores_contrast_wise.rds") %>%
  filter(contrast_reference == "oil") %>%
  transmute(tf, confidence, activity, time = str_c("Week", time, sep = " ")) %>%
  mutate(time = factor(time, levels = c("Week 8", "Week 24", "Week 48")))

mat = dorothea_scores %>%
  group_by(tf) %>%
  mutate(var = var(activity)) %>%
  mutate(mean = mean(activity)) %>%
  mutate(importance = mean) %>%
  ungroup() %>%
  filter(confidence %in% c("A", "B")) %>%
  group_by(direction = sign(importance)) %>%
  top_n(45, abs(importance)) %>% # 3 (for each time point) * 15
  ungroup() %>%
  select(-confidence, -var, -mean, -importance, -direction) %>%
  spread(time, activity) %>%
  data.frame(row.names = 1, check.names = F, stringsAsFactors = F)

color_breaks = make_div_color_breaks(mat, rwth_color("blue"), "white", rwth_color("green"))

dorothea_oil = pheatmap(mat,fontsize=12, fontsize_row = 8, 
                        color=color_breaks$colors, breaks = color_breaks$breaks, 
                        angle_col = 45, cluster_cols = F, border_color = NA,
                        cellwidth = 25) %>%
  as.ggplot()

saveRDS(dorothea_ccl4, "fig/ccl4_experiment/dorothea_oil.rds")
```

##### GO heatmap CCL4
```{r}
gsea_res = readRDS("output/ccl4_experiment/go_enrichment.rds") %>%
  filter(contrast_reference == "ccl4") %>%
  transmute(geneset, category_subcode, pval, padj, NES, time = str_c("Week", time, sep = " ")) %>%
  mutate(time = factor(time, levels = c("Week 8", "Week 24", "Week 48")),
         geneset = str_remove(geneset, "GO_"))

df = gsea_res %>%
  group_by(geneset) %>%
  mutate(mean_NES = mean(NES),
         mean_padj = mean(padj),
         var_NES = var(NES),
         importance = mean_NES) %>%
  group_by(direction = sign(importance)) %>%
  top_n(45, abs(importance)) %>%
  ungroup() %>%
  select(-c(pval, padj, mean_NES, mean_padj, var_NES, importance, direction))

anno = df %>%
  distinct(geneset, category_subcode) %>%
  data.frame(row.names = 1, check.names = F, stringsAsFactors = F)

mat = df %>%
  select(-category_subcode) %>%
  spread(time, NES) %>%
  data.frame(row.names = 1, check.names = F, stringsAsFactors = F)


color_breaks = make_div_color_breaks(mat, rwth_color("blue"), "white", rwth_color("green"))

go_ccl4 = pheatmap(mat,fontsize=12, fontsize_row = 8, 
                        color=color_breaks$colors, breaks = color_breaks$breaks, 
                        angle_col = 45, cluster_cols = F, border_color = NA,
                        cellwidth = 25, annotation_row = anno) %>%
  as.ggplot()

ggsave(go_ccl4, filename = "blub.pdf", width = 10, height = 10)
```


### Zonation Analysis
#### Overview of individual gene sets
```{r}
gene_sets = list.files("output/zonation/individual_gene_sets", pattern = "gene_sets", full.names = T) %>%
  map_df(readRDS) %>%
  filter(source != "halpern2018")

geneset_overview = gene_sets %>%
  count(source, zonation) %>%
  mutate_if(is.character, str_to_title) %>%
  ggplot(aes(x=source, y=n, fill=zonation)) +
  geom_bar(stat = "identity", position=position_dodge()) +
  labs(x = "Literature source", y="Count", fill="Zonation") +
  scale_fill_manual(values = rwth_color(c("violet", "turquoise"))) +
  my_theme("y") +
  theme(legend.position = "top")

saveRDS(geneset_overview, "fig/zonation/geneset_overview.rds")
```

#### Overlap of individual gene set
```{r}
gene_sets = list.files("output/zonation/individual_gene_sets", pattern = "gene_sets", full.names = T) %>%
  map_df(readRDS) %>%
  filter(source != "halpern2018")

# periportal
periportal = gene_sets %>% 
  filter(zonation == "periportal") %>%
  select(gene, source)

pp_h = periportal %>% filter(source == "halpern")
pp_b = periportal %>% filter(source == "braeuning")
pp_s = periportal %>% filter(source == "saito")

pp_hb = periportal %>% 
  filter(source %in% c("halpern", "braeuning")) %>% 
  group_split(source, keep = F) %>% 
  reduce(intersect)
pp_bs = periportal %>% 
  filter(source %in% c("braeuning", "saito")) %>%
  group_split(source, keep = F) %>% 
  reduce(intersect)
pp_hs = periportal %>% 
  filter(source %in% c("halpern", "saito")) %>% 
  group_split(source, keep = F) %>% 
  reduce(intersect)

pp_hbs = periportal %>% 
  group_split(source, keep = F) %>% 
  reduce(intersect)

grid.newpage()
pp_venn = draw.triple.venn(
  area1 = nrow(pp_h), area2 = nrow(pp_b), area3 = nrow(pp_s), 
  n12 = nrow(pp_hb), n23 = nrow(pp_bs), n13 = nrow(pp_hs), 
  n123 = nrow(pp_hbs),
  category = c("Halpern","Braeuning", "Saito"),
  lty = "blank",
  cex = 1,
  fontfamily = rep("sans", 7),
  fill = rwth_color(c("purple", "orange", "petrol")),
  cat.col = rwth_color(c("purple", "orange", "petrol")),
  cat.cex = 1.1,
  cat.fontfamily = rep("sans", 3)
) %>%
  as_ggplot() %>%
  grid.arrange(top = textGrob("Periportal", gp=gpar(fontsize=15, fontface = "bold")))

saveRDS(pp_venn, "fig/zonation/pp_venn.rds")

# pericentral
pericentral = gene_sets %>% 
  filter(zonation == "pericentral") %>%
  select(gene, source)

pc_h = pericentral %>% filter(source == "halpern")
pc_b = pericentral %>% filter(source == "braeuning")
pc_s = pericentral %>% filter(source == "saito")

pc_hb = pericentral %>% 
  filter(source %in% c("halpern", "braeuning")) %>% 
  group_split(source, keep = F) %>% 
  reduce(intersect)
pc_bs = pericentral %>% 
  filter(source %in% c("braeuning", "saito")) %>%
  group_split(source, keep = F) %>% 
  reduce(intersect)
pc_hs = pericentral %>% 
  filter(source %in% c("halpern", "saito")) %>% 
  group_split(source, keep = F) %>% 
  reduce(intersect)

pc_hbs = pericentral %>% 
  group_split(source, keep = F) %>% 
  reduce(intersect)

grid.newpage()
pc_venn = draw.triple.venn(
  area1 = nrow(pc_h), area2 = nrow(pc_b), area3 = nrow(pc_s), 
  n12 = nrow(pc_hb), n23 = nrow(pc_bs), n13 = nrow(pc_hs), 
  n123 = nrow(pc_hbs),
  category = c("Halpern","Braeuning", "Saito"),
  fontfamily = rep("sans", 7),
  lty = "blank",
  cex = 1,
  fill = rwth_color(c("purple", "orange", "petrol")),
  cat.col = rwth_color(c("purple", "orange", "petrol")),
  cat.cex = 1.1,
  cat.fontfamily = rep("sans", 3)
) %>% 
  as_ggplot() %>%
  grid.arrange(top = textGrob("Pericentral", gp=gpar(fontsize=15, fontface = "bold")))

saveRDS(pc_venn, "fig/zonation/pc_venn.rds")
```

#### Enrichment of consensus gene sets with CCl4 signature
```{r}
gsea_res_ccl4 = readRDS("output/zonation/consensus_and_ind_gsea_res.rds") %>%
  filter(contrast_reference == "ccl4",
         source == "consensus")

consensus_gsea_ccl4_bar = gsea_res_ccl4 %>% 
  mutate(zonation = str_to_title(zonation)) %>%
  ggplot(aes(x=time, y=NES, fill=zonation)) +
  geom_col() +
  # coord_flip() +
  scale_fill_manual(values = rwth_color(c("violet", "turquoise"))) +
  labs(x = "Time in weeks", fill = "Zonation") +
  theme(legend.position = "top") +
  geom_text(aes(label = str_c("FDR:", scientific(padj,3), sep = " "), y=0.95*NES), color="white", position=position_dodge(width=0), size=4.5) +
  my_theme(grid = "y")

saveRDS(consensus_gsea_ccl4_bar, "fig/zonation/consensus_pc_pp_gsea_ccl4.rds")
```

#### Enrichment of consensus gene sets with Oil signature
```{r}
gsea_res_oil = readRDS("output/zonation/consensus_and_ind_gsea_res.rds") %>%
  filter(contrast_reference == "oil",
         source == "consensus")

consensus_gsea_oil_bar = gsea_res_oil %>% 
  mutate(zonation = str_to_title(zonation)) %>%
  ggplot(aes(x=time, y=NES, fill=zonation)) +
  geom_col() +
  # coord_flip() +
  scale_fill_manual(values = rwth_color(c("violet", "turquoise"))) +
  labs(x = "Time in weeks", fill = "Zonation") +
  theme(legend.position = "top") +
  geom_text(aes(label = str_c("FDR:", scientific(padj,3), sep = " "), y=0.95*NES), color="white", position=position_dodge(width=0), size=4.5) +
  my_theme(grid = "y")

saveRDS(consensus_gsea_oil_bar, "fig/zonation/consensus_pc_pp_gsea_oil.rds")
```

#### GSEA plots CCL4
```{r}
gsea_res_ccl4 = readRDS("output/zonation/consensus_and_ind_gsea_res.rds") %>%
  filter(contrast_reference == "ccl4",
         source == "consensus")

zonation_gene_sets = readRDS("output/zonation/consensus_gene_sets.rds") %>%
  group_by(zonation) %>%
  summarise(gene_set = list(gene)) %>%
  deframe()

gsea_plots_ccl4 = gsea_res_ccl4 %>% 
  inner_join(enframe(zonation_gene_sets, "zonation", "geneset"), by="zonation") %>%
  mutate(gsea_plots = pmap(., .f = make_gsea_plot))

saveRDS(gsea_plots_ccl4, "fig/zonation/gsea_plots_ccl4.rds")
```

#### GSEA plots of Oil
```{r}
gsea_res_oil = readRDS("output/zonation/consensus_and_ind_gsea_res.rds") %>%
  filter(contrast_reference == "oil",
         source == "consensus")

zonation_gene_sets = readRDS("output/zonation/consensus_gene_sets.rds") %>%
  group_by(zonation) %>%
  summarise(gene_set = list(gene)) %>%
  deframe()

gsea_plots_oil = gsea_res_oil %>% 
  inner_join(enframe(zonation_gene_sets, "zonation", "geneset"), by="zonation") %>%
  mutate(gsea_plots = pmap(., .f = make_gsea_plot))

saveRDS(gsea_plots_oil, "fig/zonation/gsea_plots_oil.rds")
```

#### Enrichment of individual gene sets with ccl4 signature
```{r}
gsea_res_ccl4_ind = readRDS("output/zonation/consensus_and_ind_gsea_res.rds")%>%
  filter(contrast_reference == "ccl4",
         source != "consensus")

ind_gsea_ccl4_bar = gsea_res_ccl4_ind %>% 
  mutate(zonation = str_to_title(zonation),
         source = str_to_title(source)) %>%
  ggplot(aes(x=time, y=NES, fill=zonation)) +
  geom_col(position=position_dodge()) +
  facet_rep_wrap(~source) +
  scale_fill_manual(values = rwth_color(c("violet", "turquoise"))) +
  labs(x = "Time in weeks", fill = "Zonation") +
  theme(legend.position = "top") +
  # geom_text(aes(label = str_c("FDR:", scientific(padj,3), sep = " "), y=0.75*NES), color="white", position=position_dodge(width=0.9), size=4, angle = -90) +
  geom_text(aes(label = gtools::stars.pval(padj), y=0.9*NES), color="white", position=position_dodge(width=0.9), size=5) +
  my_theme("y")

saveRDS(ind_gsea_ccl4_bar, "fig/zonation/ind_pc_pp_gsea_ccl4.rds")
```

#### Enrichment of individual gene sets with oil signature
```{r}
gsea_res_oil_ind = readRDS("output/zonation/consensus_and_ind_gsea_res.rds")%>%
  filter(contrast_reference == "oil",
         source != "consensus")

ind_gsea_oil_bar = gsea_res_oil_ind %>% 
  mutate(zonation = str_to_title(zonation),
         source = str_to_title(source)) %>%
  ggplot(aes(x=time, y=NES, fill=zonation)) +
  geom_col(position=position_dodge()) +
  facet_rep_wrap(~source) +
  scale_fill_manual(values = rwth_color(c("violet", "turquoise"))) +
  labs(x = "Time in weeks", fill = "Zonation") +
  theme(legend.position = "top") +
  # geom_text(aes(label = str_c("FDR:", scientific(padj,3), sep = " "), y=0.75*NES), color="white", position=position_dodge(width=0.9), size=4, angle = -90) +
  geom_text(aes(label = gtools::stars.pval(padj), y=0.9*NES), color="white", position=position_dodge(width=0.9), size=5) +
  my_theme("y")

saveRDS(ind_gsea_oil_bar, "fig/zonation/ind_pc_pp_gsea_oil.rds")
```

