---
title: "Zonation analysis"
author: "Christian Holland"
date: "07/02/2019"
output: html_document
---

```{r "knitr config", cache=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::knit(..., quiet = TRUE)
```

## Libraries and sources
These libraries and sources are used in this analysis 
```{r "setup", message=F}
library(scran)
library(tidyverse)
library(tidylog)
library(janitor)
library(readxl)
library(limma)
library(Matrix)
library(biobroom)
library(clinfun)
library(fgsea)
library(msigdf)

options("tidylog.display" = list(print))
```

## Zonation: Periportal vs pericentral genes
### Process external data
#### Halpern et al.
##### Working with raw data and assignment of cells manually
###### Build zonation matrix
```{r}
df = read_excel("data/zonation/Halpern/zonation_matrix.xlsx", skip = 1) %>%
  clean_names() %>%
  rename(cell = cell_number) %>%
  mutate(cell = str_to_lower(str_replace(cell, " ", "_"))) %>%
  gather(layer, prob, -cell) %>%
  mutate(layer = ordered(parse_number(layer)))

cell_zonation_map = df %>% 
  group_by(cell) %>%
  slice(which.max(prob)) %>%
  ungroup()

saveRDS(cell_zonation_map, "output/zonation/halpern/cell_zonation.rds")
```

###### Process and normalize count matrix
```{r}
count_matrix = read_csv("data/zonation/Halpern/count_matrix.csv") %>%
  data.frame(row.names = 1) %>%
  as.matrix() %>%
  Matrix(sparse=T)

# how oft a gene is expressed?
apply(count_matrix,1,function(row) {
  sum(row != 0)
}
) %>%
  enframe() %>%
  ggplot(aes(x=log10(value+1))) +
  geom_density() +
  geom_vline(xintercept = log10(15+1))

# keep only genes that are expressed in more than 15 out of 1415 cells
keep_genes = apply(count_matrix,1, function(row) {
      sum(row != 0) >= 15
})

norm = SingleCellExperiment(list(counts=count_matrix[keep_genes, ])) %>%
  computeSumFactors() %>%
  normalize() %>%
  exprs() 

saveRDS(norm, "output/zonation/halpern/norm_expression.rds")
```

###### Construct gene sets
```{r}
norm = readRDS("output/zonation/halpern/norm_expression.rds")
cell_zonation = readRDS("output/zonation/halpern/cell_zonation.rds")

x = norm %>% 
  as.matrix() %>%
  data.frame() %>%
  rownames_to_column("gene") %>%
  as_tibble() %>%
  gather(cell, expression, -gene) %>%
  left_join(cell_zonation, by="cell")

# summarizing expression per gene and layer
summarized_scores = x %>%
  group_by(gene, layer) %>%
  summarise(mean_expression = mean(expression)) %>%
  ungroup() %>%
  separate_rows(gene, sep = ";")

saveRDS(summarized_scores, "output/zonation/halpern/summarized_scores.rds")

stats = x %>%
  nest(-gene) %>%
  mutate(stat = data %>% map(function(data) {
    # Testing that expression changes monotonic across the layer axis
    jonckheere.test(data$expression, data$layer, alternative = "two.sided") %>%
      broom::tidy()
  })) %>%
  mutate(direction = data %>% map(function(data) {
    # exploiting the slope of linear models to identify whether the genes 
    # expressed increase or decrease monotonically
    lm(expression~as.integer(layer), data=data) %>%
      broom::tidy() %>%
      filter(term == "as.integer(layer)") %>%
      pull(estimate)
  })) %>%
  select(-data) %>%
  unnest(stat) %>%
  unnest(direction) %>%
  mutate(fdr = p.adjust(p.value, method = "BH"))

saveRDS(stats, "output/zonation/halpern/jonck_test.rds")

mgi_symbols = read_csv("data/annotation/mgi_ensembl_entrez.csv") %>%
  drop_na(mgi_symbol) %>%
  distinct(gene = mgi_symbol)


halpern = stats %>%
  filter(fdr <= 0.001) %>%
  mutate(zonation = case_when(sign(direction) == 1 ~ "periportal",
                              sign(direction) == -1 ~ "pericentral")) %>%
  mutate(source = "halpern") %>%
  select(gene, zonation, source) %>%
  separate_rows(gene, sep = ";") %>%
  mutate(gene = alias2SymbolTable(gene, species="Mm")) %>%
  drop_na() %>%
  distinct() %>%
  semi_join(mgi_symbols, by="gene")

saveRDS(halpern, "output/zonation/individual_gene_sets/halpern_gene_sets.rds")

```

#### Braeuning et al
```{r}
mgi_symbols = read_csv("data/annotation/mgi_ensembl_entrez.csv") %>%
  drop_na(mgi_symbol) %>%
  distinct(gene = mgi_symbol)

braeuning = read_excel(
  "data/zonation/Braeuning/Brauning et al_TableS1_log2ratios.xlsx",
  col_names = c("accession", "gene_name", "gene", "logfc"), skip = 1
  ) %>%
  drop_na() %>%
  # remove text artefacts (\r and \n)
  mutate(logfc = str_remove(logfc, "\\\r")) %>%
  mutate(logfc = str_remove(logfc, "\\\n")) %>%
  # gene represented by more than one probe are separated
  separate_rows(logfc, sep="/", convert = T) %>%
  # remove text artefacts (\r and \n)
  mutate(gene = str_remove(gene, "\\\r")) %>%
  mutate(gene = str_remove(gene, "\\\n")) %>%
  # genes summarized into a single term are seprated (e.g. Raet1a/b/c/d/e)
  separate_rows(gene, sep ="///") %>%
  mutate(gene = str_trim(gene)) %>%
  # update gene symbol
  mutate(gene = alias2SymbolTable(gene, species = "Mm")) %>%
  drop_na(gene) %>%
  # consider only official mgi symbols
  semi_join(mgi_symbols) %>%
  group_by(gene) %>%
  summarise(logfc = mean(logfc)) %>%
  mutate(source = "braeuning",
         zonation = case_when(sign(logfc) == 1 ~ "pericentral",
                              sign(logfc) == -1 ~ "periportal"))

saveRDS(braeuning, 
        "output/zonation/individual_gene_sets/braeuning_gene_sets.rds")
```

#### Saito et al
```{r}
mgi_symbols = read_csv("data/annotation/mgi_ensembl_entrez.csv") %>%
  drop_na(mgi_symbol) %>%
  distinct(gene = mgi_symbol)

saito_pericentral_male = read_excel(
  "data/zonation/Saito/Copy of male zone 3 specific high GO analysis for table (formatted).xls",
  col_names = c("biological_process", "gene", "alternative_gene_symbol", 
                "gene_name", "logfc")
  )

saito_pericentral_common = read_excel(
  "data/zonation/Saito/Copy of zone 3 common high GO analysis for table (formatted).xls",
  col_names = c("biological_process", "gene", "alternative_gene_symbol", 
                "gene_name", "logfc")
)

saito_pericentral = bind_rows(saito_pericentral_male, 
                              saito_pericentral_common) %>%
  drop_na(gene) %>%
  mutate(gene = alias2SymbolTable(gene, species = "Mm")) %>%
  drop_na(gene) %>%
  semi_join(mgi_symbols) %>%
  group_by(gene) %>%
  summarise(logfc = mean(logfc)) %>%
  mutate(source = "saito",
         zonation = "pericentral")

saito_periportal_male = read_excel(
  "data/zonation/Saito/Copy of male zone 1 specific high GO analysis for table (formatted).xls",
  col_names = c("biological_process", "gene", "alternative_gene_symbol", "gene_name", "logfc")
  ) 

saito_periportal_common = read_excel(
  "data/zonation/Saito/Copy of zone 1 common high GO analysis for table (formatted).xls",
  col_names = c("biological_process", "gene", "alternative_gene_symbol", "gene_name", "logfc")
)

saito_periportal = bind_rows(saito_periportal_male, saito_periportal_common) %>%
  drop_na(gene) %>%
  mutate(gene = alias2SymbolTable(gene, species = "Mm")) %>%
  drop_na(gene) %>%
  semi_join(mgi_symbols) %>%
  group_by(gene) %>%
  summarise(logfc = mean(logfc)) %>%
  mutate(source = "saito",
         zonation = "periportal")

saito = bind_rows(saito_pericentral, saito_periportal)

saveRDS(saito, "output/zonation/individual_gene_sets/saito_gene_sets.rds")
```

### Analyse gene sets
#### Construct consensus periportal and pericentral gene sets
```{r}
# We consider genes as gene set member when they appear at least 2 times
gene_sets = list.files("output/zonation/individual_gene_sets", 
                       pattern = "gene_sets", full.names = T) %>%
  map_df(readRDS) %>%
  filter(source != "halpern2018") %>%
  count(gene, zonation) %>%
  filter(n>=2) %>%
  distinct(gene, zonation) %>%
  arrange(zonation, gene)

saveRDS(gene_sets, "output/zonation/consensus_gene_sets.rds")
```

#### Gene set enrichment analysis with consensus and individuala pc/pp gene et and ccl4/oil signatures
```{r}
consensus_zonation_gene_sets = readRDS(
  "output/zonation/consensus_gene_sets.rds"
  ) %>%
  mutate(source = "consensus")

individual_zonation_gene_sets = list.files(
  "output/zonation/individual_gene_sets", pattern = "gene_sets", full.names = T
  ) %>%
  map_df(readRDS) %>%
  filter(source != "halpern2018") 

zonation_gene_sets = bind_rows(consensus_zonation_gene_sets, 
                               individual_zonation_gene_sets) %>%
  unite(source, source, zonation) %>%
  group_by(source) %>%
  summarise(gene_set = list(gene)) %>%
  deframe()

limma_result = readRDS("output/ccl4_experiment/limma_result.rds") %>%
  filter(contrast_reference %in% c("ccl4", "oil", "treatment")) %>%
  mutate(time = str_extract(contrast, "\\d*$"),
         time = ordered(as.integer(time)))

# GSEA
set.seed(123)
gsea_res = limma_result %>%
  nest(-c(contrast, contrast_reference, time)) %>%
  mutate(gsea = data %>% map(function(data) {
    stats = data %>% 
      select(gene,statistic) %>%
      deframe()
    
    fgsea(pathways = zonation_gene_sets, stats = stats, nperm = 10000) %>%
      as_tibble() %>%
      separate(pathway, into = c("source", "zonation"))
  })) %>%
  unnest(gsea, .preserve = data)

saveRDS(gsea_res, "output/zonation/consensus_and_ind_gsea_res.rds")
```

#### ORA to characterize consensus gene sets
```{r}
dorothea_genesets = read_csv(
  "https://raw.githubusercontent.com/saezlab/ConservedFootprints/master/data/dorothea_benchmark/regulons/dorothea_regulon_mouse_v1.csv"
) %>%
  select(-mor, -likelihood) %>%
  rename(geneset = tf, gene = target) %>%
  mutate(category_code = "tf",
         category_subcode = "dorothea") %>%
  nest(gene, .key="genes")

progeny_genesets = read_csv(
  "https://raw.githubusercontent.com/saezlab/ConservedFootprints/master/data/progeny_benchmark/models/progeny_matrix_mouse_v1.csv"
  ) %>%
  rename(gene = X1) %>%
  gather(pathway, weight, -gene) %>%
  filter(weight != 0) %>%
  select(-weight) %>%
  rename(geneset = pathway) %>%
  mutate(category_code = "pathway",
         category_subcode = "progeny") %>%
  nest(gene, .key="genes")
  
go_genesets = msigdf.mouse %>%
  filter(category_code == "c5") %>%
  filter(category_subcode == "bp" | category_subcode == "mf") %>%
  rename(gene = mouse.symbol) %>%
  nest(gene, .key="genes") 

kegg_genesets = msigdf.mouse %>%
  filter(category_code == "c2") %>%
  filter(str_detect(geneset, "KEGG_")) %>%
  rename(gene = mouse.symbol) %>%
  nest(gene, .key="genes") 
  
genesets = bind_rows(dorothea_genesets, progeny_genesets, go_genesets, kegg_genesets)


conf_matrices = readRDS("output/zonation/consensus_gene_sets.rds") %>%
  nest(-zonation, .key="signature") %>%
  mutate(genesets = list(genesets)) %>%
  mutate(conf_matrices = pmap(., .f=function(signature, genesets, zonation...) {
    # set = genesets %>% pluck(2,1)
    genesets %>%
      mutate(conf_matrix = genes %>% map(function(set) {
        
        # "signature" is either the periportal or pericentral gene set
        # "set" is a classical gene set like GO-term
        
        matrix(
          c(
            nrow(intersect(set, signature)), nrow(setdiff(set, signature)),
            nrow(setdiff(signature, set)), 20000 - nrow(signature) - nrow(setdiff(set, signature))
          ), 
          nrow = 2, ncol=2, byrow = T, 
          dimnames = list("Geneset" = c("Member", "Non-Member"),
                          "PC/PP-Geneset" =  c("Member", "Non-Member"))
        )
      })) %>%
      select(-genes)
  })) %>%
  unnest(conf_matrices)

ora_result = conf_matrices %>%
  mutate(ora = conf_matrix %>% map(~tidy(fisher.test(.)))) %>%
  unnest(ora) %>%
  group_by(zonation, category_subcode) %>%
  mutate(qval = p.adjust(p.value, method = "bonferroni"),
         fdr = p.adjust(p.value, method = "BH")) %>%
  ungroup()

saveRDS(ora_result, "output/zonation/ora_result_of_consensus_genesets.rds")
```

#### Overlap between CCL4/Oil signatures and pericentral and periportal genesets
```{r}
zonation_gene_sets = readRDS("output/zonation/consensus_gene_sets.rds")
pp = zonation_gene_sets %>% filter(zonation == "periportal") %>% pull(gene)
pc = zonation_gene_sets %>% filter(zonation == "pericentral") %>% pull(gene)

# relax requirement for significant genes (originally effect size 1.5 and fdr 0.05)
limma_result = readRDS("output/ccl4_experiment/limma_result.rds") %>%
  filter(contrast_reference %in% c("ccl4", "oil")) %>%
  mutate(regulation = case_when(logFC >= 0.8 & fdr <= 0.2 ~ "up",
                                logFC <= -0.8 & fdr <= 0.2 ~ "down",
                                TRUE ~ "no")) %>%
  filter(regulation != "no") %>%
  mutate(time = str_extract(contrast, "\\d*$"),
         time = ordered(as.integer(time)))

# data = overlap_coef_df %>% pluck(5,1)
overlap_coef_df = limma_result %>%
  nest(-c(contrast, contrast_reference, time, regulation)) %>%
  mutate(overlap = data %>% map(function(data) {
    signature = data %>% pull(gene)
    
    # extract common genes
    pp_common_genes = intersect(pp, signature)
    pc_common_genes = intersect(pc, signature)
    
    # calculate overlap coefficient
    pp_overlap = length(pp_common_genes) / min(length(pp), length(signature))
    pc_overlap = length(pc_common_genes) / min(length(pc), length(signature))
    
    # summarize results in a tibble
    overlap_df = tibble(zonation = c("periportal", "pericentral"),
                        overlap_coef = c(pp_overlap, pc_overlap),
                        n_common = c(length(pp_common_genes), length(pc_common_genes)),
                        overlap = list(pp_common_genes, pc_common_genes))
    
  }))

saveRDS(overlap_coef_df, "output/zonation/overlap_ccl4_oil_sig_with_pc_pp_genesets.rds")
```

#### ORA to characterize overlap of ccl4 and oil signature and pericentral and periportal genesets
```{r}
signature = readRDS("output/zonation/overlap_ccl4_oil_sig_with_pc_pp_genesets.rds") %>%
  unnest(overlap) %>%
  unnest(overlap) %>%
  # distinct(regulation, time, zonation, overlap) %>%
  distinct(contrast_reference, regulation, zonation, overlap) %>%
  rename(gene = overlap) %>%
  nest(gene, .key="signature")


dorothea_genesets = read_csv(
  "https://raw.githubusercontent.com/saezlab/ConservedFootprints/master/data/dorothea_benchmark/regulons/dorothea_regulon_mouse_v1.csv"
) %>%
  select(-mor, -likelihood) %>%
  rename(geneset = tf, gene = target) %>%
  mutate(category_code = "tf",
         category_subcode = "dorothea") %>%
  nest(gene, .key="genes")

progeny_genesets = read_csv(
  "https://raw.githubusercontent.com/saezlab/ConservedFootprints/master/data/progeny_benchmark/models/progeny_matrix_mouse_v1.csv"
  ) %>%
  rename(gene = X1) %>%
  gather(pathway, weight, -gene) %>%
  filter(weight != 0) %>%
  select(-weight) %>%
  rename(geneset = pathway) %>%
  mutate(category_code = "pathway",
         category_subcode = "progeny") %>%
  nest(gene, .key="genes")
  

go_genesets = msigdf.mouse %>%
  filter(category_code == "c5") %>%
  filter(category_subcode == "bp" | category_subcode == "mf") %>%
  rename(gene = mouse.symbol) %>%
  nest(gene, .key="genes") 

kegg_genesets = msigdf.mouse %>%
  filter(category_code == "c2") %>%
  filter(str_detect(geneset, "KEGG_")) %>%
  rename(gene = mouse.symbol) %>%
  nest(gene, .key="genes") 
  
genesets = bind_rows(dorothea_genesets, progeny_genesets, go_genesets, kegg_genesets)

# distribution of geneset size per category_subcode
genesets %>% 
  transmute(category_subcode, geneset, size = genes %>% map(nrow)) %>%
  unnest(size) %>%
  arrange(-size)
  ggplot(aes(x = size)) +
  geom_histogram() +
  facet_wrap(~category_subcode, scales="free")

# signature = conf_matrices %>% pluck(3,1)
# genesets = conf_matrices %>% pluck(4,1)
conf_matrices = signature %>%
  mutate(genesets = list(genesets)) %>%
  mutate(conf_matrices = pmap(., .f=function(signature, genesets, ...) {
    set = genesets %>% pluck(2,1)
    genesets %>%
      mutate(conf_matrix = genes %>% map(function(set) {
        
        # "signature" is either the overlap of deg from oil and ccl4 and periportal or pericentral gene set
        # "set" is a classical gene set like GO-term
        
        matrix(
          c(
            nrow(intersect(set, signature)), nrow(setdiff(set, signature)),
            nrow(setdiff(signature, set)), 20000 - nrow(signature) - nrow(setdiff(set, signature))
          ), 
          nrow = 2, ncol=2, byrow = T, 
          dimnames = list("Gene-Set" = c("Member", "Non-Member"),
                          "Gene-Signature" =  c("Member", "Non-Member"))
        )
      })) %>%
      select(-genes)
  })) %>%
  unnest(conf_matrices)

ora_result = conf_matrices %>%
  mutate(ora = conf_matrix %>% map(~tidy(fisher.test(.)))) %>%
  unnest(ora) %>%
  group_by(contrast_reference, regulation, zonation, category_subcode) %>%
  mutate(qval = p.adjust(p.value, method = "bonferroni"),
         fdr = p.adjust(p.value, method = "BH")) %>%
  ungroup()

saveRDS(ora_result, "output/zonation/ora_result_of_overlap.rds")
```